Finding and controling biases in dark matter halo concentrations estimates

### 2016-07-05 paper/ms.tex
restructure conclusions

Avoiding biases with varying simulation resolution is one of the most
important goals of a robust method to estimate halo concentration. 
We propose bootstraping individual high resolution halos to quantify these
possible biases.
In this experiment we take halos sampled with at least $10^{5}$
particles and subsample each halo by factors of $2$ up to $10^{3}$ to
measure the concentration every time the halo is resampled.

@@

In this paper we presented a new method to estimate the concentration
of dark matter halos in N-body simulations.  We tested the method on
mock halo data to study the impact of total number of particles and
input concentration on the retrieved values.  We compared these
results against two other methods commonly used in the literature to
estimate concentrations.  For all methods, the accuracy in retrieving
the input concentration increases with the number of particles as
summarized in Figure \ref{fig:error}.  The new method systematically
outperforms the other two.

We also applied the method to halos extracted from a cosmological
N-body simulation.  Using this data we showed that the new method is
robust with respect to a decreasing number of particles in the halos.
On the other hand we showed that the velocity method has a systematic
bias that overestimates the concentration for halos with a low particle
number up to a factor of $25\%$.

We provide for the first time in the literature a direct way to
measure the bias in the estimated halo concentration at low particle numbers and
a way to use it to make corrections on the estimates. 
In Section \ref{subsec:cosmo} we presented a simple method that can be used on any
dataset to estimate the errors and correct them when estimating the
concentration at low particle numbers.

>>>>

We use bootstraping of individual high resolution halos to quantify
biases for different methods used to esteimate concentrations.

In this experiment we take halos sampled with at least $10^{5}$
particles and subsample each halo by factors of $2$ up to $10^{3}$ to
measure the concentration every time the halo is resampled.

@@

In this paper we used bootstrapping on dark matter halos resolved with a
large number of particles ($\sim 10^5$) in order to quantify biases on
concentration estimates. 
We found that methods commonly used in the literature present a large
bias in the concentration by overestimating it by factors of
$15\%-20\%$ for halos with $200$ particles. 
The scaling of the bias as a function of particle number was found in
two very different kinds of simulations, one cosmological and the
other of an individual high resolution halo, suggesting its universal
nature.

This provided a motivation to introduce a new method to estimate the
concentration that shows a lower bias.
The method is based on the integrated mass profile. 
We also implemented a fitting procedure based on a Markov Chain Monte Carlo
algorithm.
In the bootstrapping experiment this method showed a bias on the order
of $2\%-3\%$ for halos with $200$ particles.

We provide for the first time in the literature a direct way to
measure the bias in the estimated halo concentration at low particle numbers and
a way to use it to make corrections on the estimates.
